---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF"
author: "Maresa Behr (11119838)"
date: "`r Sys.Date()`"
output: distill::distill_article
---

## General Remarks

-   You can submit your solutions in teams of up to 3 students.
-   Include all your team-member's names and student numbers (Martrikelnummern) in the `authors` field.
-   Please use the exercise template document to work on and submit your results.
-   Use a level 2 heading for each new exercise and answer each subtask next to it's bullet point or use a new level 3 heading if you want.
-   Always render the R code for your solutions and make sure to include the resulting data in your rendered document.
    -   Make sure to not print more than 10 rows of data (unless specifically instructed to).
-   Always submit both the rendered document(s) as well as your source Rmarkdown document. Submit the files separately on moodle, **not** as a zip archive.

## Exercise 1: Initializing git (4 Points)

For this whole exercise sheet we will be tracking all our changes to it in git.

a)  Start by initializing a new R project with git support, called `2024-exeRcise-sheet-3`. If you forgot how to do this, you can follow this [guide](https://malikaihle.github.io/Introduction-RStudio-Git-GitHub/rstudio_project.html).
b)  Commit the files generated by Rstudio.
c)  For all of the following tasks in this exercise sheet we ask you to always commit your changes after finishing each subtask e.g. create a commit after task *1d*, *1e* etc.

> Note: This applies only to answers that have text or code as their answer. If you complete tasks in a different order or forget to commit one, this is no problem. If you change your answers you can just create multiple commits to track the changes.

d)  Name 2 strengths and 2 weaknesses of git. (Don't forget to create a commit after this answer, see *1c*)

    strength 1: it makes it easy to collaborate on a project with a team

    strenght 2: it allows you to track the changes you've made, which helps to resolve problems

    weakness 1: when working with binary files problems will occur

    weakness 2: it doesn't run on its own in the background, you have to explicitly push changes to it

e)  Knit this exercise sheet. Some new files will automatically be generated when knitting the sheet e.g. the HTML page. Ignore these files, as we only want to track the source files themselves.

## Exercise 2: Putting your Repository on GitHub (3.5 Points)

For this task you will upload your solution to GitHub.

a)  Create a new repository on GitHub in your account named `exeRcise-sheet-3`. Make sure you create a **public repository** so we are able to see it for grading. Add the link to the repository below:

    <https://github.com/maresamika/exeRcise-sheet-3>

b)  Push your code to this new repository by copying and executing the snippet on github listed under `…or push an existing repository from the command line`.

c)  Regularly push your latest changes to GitHub again and especially do so when you are finished with this sheet.

## Exercise 3: Baby-Names in Munich (3.5 Points)

Download the latest open datasets on given names ("Vornamen") from the open data repository of the city of Munich for the years `2023` and `2022`.

Link: <https://opendata.muenchen.de/dataset/vornamen-von-neugeborenen>

a)  Download the data for both years and track it in git. For small datasets like these adding them to git is not a problem.

b)  Load the data for both years into R. Check the type of the count variable ("Anzahl") and look into the data to determine why it is not numeric? Fix the problem in an appropriate manner, it is OK if some of the counts are inaccurate because of this. Explain your solution and the repercussions.

    For both datasets I choose to mutate the values of the column Anzahl, that are matching the string "4 oder weniger" to 4. Then I set the column to numeric.

    Names that were counted 1,2 or 3 times are now tipping the scales as they are more weighted by being set to 4 instead of the correct amount. Since we don't know the exact numbers in the first place, the repercussions for working with this data are rather little.

    ```{r}
    library(tidyverse)

    vornamen_2022 <- read.csv(file.path('data', 'open_data_portal_2022.csv'))
    vornamen_2023 <- read.csv(file.path('data', 'vornamen-muenchen-2023.csv'))

    #Anzahl is not numeric because of value "4 oder weniger"
    vornamen_2022 <- vornamen_2022 %>% 
      mutate(Anzahl = str_replace_all(Anzahl, "4 oder weniger", "4")) %>% 
      mutate(Anzahl = as.numeric(Anzahl))
    #same for vornamen_2023
    vornamen_2023 <- vornamen_2023 %>% 
      mutate(Anzahl = str_replace_all(Anzahl, "4 oder weniger", "4")) %>% 
      mutate(Anzahl = as.numeric(Anzahl))


    ```

c)  Calculate the total number of babies born in Munich in 2022 and 2023. Which year had the bigger baby-boom?

    the year 2022 had the bigger baby boom.

    ```{r}
    sum(vornamen_2022$Anzahl)
    sum(vornamen_2023$Anzahl)
    ```

d)  Add a new column `year` to both datasets which holds the correct year for each.

    ```{r}
    vornamen_2022 <- vornamen_2022 %>% 
      mutate(year = 2022)
    vornamen_2023 <- vornamen_2023 %>% 
      mutate(year = 2023)
    ```

e)  Combine both datasets into one using `bind_rows()`.

    ```{r}
    vornamen_22_23 <- bind_rows(vornamen_2022, vornamen_2023)
    ```

f)  Combine the counts for same names to determine the most popular names across both years. Print out the top 10 names in a nicely formatted table for both years. Include a table caption.

    ```{r}
    vornamen_22_23 %>% 
      group_by(Vorname) %>% 
      summarize(anzahl_gesamt = sum(Anzahl)) %>% 
      arrange(desc(anzahl_gesamt)) %>% 
      head(10) %>% 
      knitr::kable(., 
                   caption = "Die 10 beliebtesten Vornamen in 2022 und 2023" ,
                   col.names = c("Vorname", "Gesamtanzahl")) %>% 
      kableExtra::kable_minimal()
    ```

## Exercise 4: Open Analysis (4 points)

This exercise is a bit more open-ended. You can choose any dataset from [Our World in Data](https://ourworldindata.org/) and analyze it, while determining the research question yourself.

a)  Go to <https://github.com/owid/owid-datasets/tree/master/datasets> and choose a dataset that interests you. You can have a look at <https://ourworldindata.org/> to gather some inspiration.

b)  Download the dataset and track it in git.

    ```{r}
    natural_disasters <- read.csv(file.path('data', 'Natural disasters (EMDAT).csv'))
    ```

c)  Put the name / title of the dataset and a link to it below.

-   Dataset Name: Natural Disasters (EMDAT)

-   Link: [https://github.com/owid/owid-datasets/tree/master/datasets/Natural%20disasters%20(EMDAT)](https://github.com/owid/owid-datasets/tree/master/datasets/Natural%20disasters%20(EMDAT)){.uri}

d)  Come up with a (research) question you want to answer with the data and briefly explain why you believe this is an interesting question within one sentence. It should be a question that can be answered with the dataset and using R.
e)  Use R to answer your chosen question.
f)  Create a meaningful plot / figure with the dataset. Make sure to provide a figure caption (via the chunk options / Rmarkdown) and correctly label the figure.

## Final Note

Make sure to push all your commits and changes to GitHub before submittining the exercise sheet.
